---
title: "Preliminary look at possible latent variables, YL-YC"
output:
  pdf_document: default
  html_notebook: default
---

We take a look at possible latent variables, working backwards from outcomes

First, we load the data file, and some packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load some packages

```{r load_packages}

library(pacman) #install this if not already present
p_load(tidyverse, haven, magrittr, psych, car, knitr, lavaan, GPArotation, reshape2)
```

Then we clear the workspace, and load the data file created in the script "import data from YL files script".

```{r load_data}

rm(list=ls(all=TRUE))

load("peru_long_yc.Rdata")

```



# 1  Overall measurement model 

Let's try putting a measurement model together for the whole model

```{r overall+measure_model}


temp <- peru_long.dat %>% 
          select(childid, round, typesite, indigenous,
                 mother_age_birth, partner, momedu, dadlit, momlit, 
                 wi_early_yrs, food_insecurity_r2, totalexp,
                 chore_hours, school_type, hours_study, read_encourage,
                 sdq1, sdq3, sdq4, sdq5, 
                 sdq9, sdq8, sdq10, sdq11, 
                 sdq13, sdq16, sdq12, sdq14,
                 sdq17, sdq20, sdq15, sdq18, 
                 sdq19,
                 drink_to_drunk, drunk_hit, alc_1perwk,
                 stunting_to_r3, probs_vision, probs_resp,
                 care_school_qual2, care_school_qual3,
                 care_school_qual4, 
                 school_punish1, school_punish2,
                 lang_instruct_matches,
                 ppvt_raw, maths_perco, yrs_grtr_grade, literate, deprived_grouping) %>% 
          melt(id.vars = c("childid","round")) %>% 
          dcast(childid ~ round + variable)
            

temp %<>% 
  mutate(`1_mother_age_birth` = `1_mother_age_birth`/25,
         `2_totalexp` = `2_totalexp`/750,
         `4_ppvt_raw`= `4_ppvt_raw`/200,
         `4_maths_perco`= `4_maths_perco`/200,
         `3_chore_hours`= `3_chore_hours`/2.2,
         `3_read_encourage` = `3_read_encourage`/3.2,
         `3_stunting_to_r3` = `3_stunting_to_r3`/2)

temp %<>% 
  select(`1_mother_age_birth`, `1_partner`,
                   `1_momedu`, `1_dadlit`, `1_momlit`, `1_typesite`, `1_indigenous`,

                   `2_wi_early_yrs`, `2_food_insecurity_r2`, `2_totalexp`,

                   `3_chore_hours`, `3_school_type`, `3_hours_study`,
                                `3_read_encourage`,

                             `1_sdq1`, `1_sdq3`, `1_sdq4`, `1_sdq5`,  
                               `1_sdq9`, `1_sdq8`, `1_sdq10`, `1_sdq11`,  
                               `1_sdq13`, `1_sdq14`, 
                               `1_sdq17`, `1_sdq20`, `1_sdq15`, `1_sdq18`, 
                    `1_sdq19`,
                    `1_drink_to_drunk`, `1_drunk_hit`, `1_alc_1perwk`,


                # Child health
                    `3_stunting_to_r3`, `3_probs_vision`, `3_probs_resp`,

                # School environment
                    `3_care_school_qual2`, `3_care_school_qual3`, 
                    `3_care_school_qual4`, 
                    `3_school_punish1`, `3_school_punish2`,  
                    `3_lang_instruct_matches`,

                # Round`3 outcomes
                    `3_ppvt_raw`, `3_maths_perco`, `3_yrs_grtr_grade`, 
                                  `3_literate`, 
                # Round`4 outcomes
                   `4_ppvt_raw`, `4_maths_perco`, `4_yrs_grtr_grade`) 

# summary(temp)

temp.model <- '
               # Caregiver Status
                   care_status =~ `1_partner` + `1_mother_age_birth`
                   lit_ed =~`1_momedu` +`1_momlit`
                   

               # # Economic well-being
                    eco_well_being_cfa =~`2_wi_early_yrs` +`2_food_insecurity_r2` +`2_totalexp`

               # Education investment
                   ed_invest =~`3_chore_hours` +`3_school_type` +`3_hours_study`+
                                `3_read_encourage`
         

               # Adverse EC environment
                    mmhealth =~`1_sdq1` +`1_sdq3` +`1_sdq4` +`1_sdq5` + 
                               `1_sdq9` +`1_sdq8` +`1_sdq10` +`1_sdq11` + 
                               `1_sdq13` +`1_sdq14` +
                               `1_sdq17 ` +`1_sdq20` +`1_sdq15` +`1_sdq18` +
                               `1_sdq19`
                    
                # Child health
                    child_health =~`3_stunting_to_r3` +`3_probs_vision` +`3_probs_resp`

                # School environment
                   #  school_qual =~`3_care_school_qual2` +`3_care_school_qual3` +
                   # `3_care_school_qual4` 
                    # school_punish =~`3_school_punish1` +`3_school_punish2`  
                    lang_instruct =~`3_lang_instruct_matches`
                     # school_environ =~ school_qual + lang_instruct
                     # school_environ ~~ 1*school_environ

                # Round`3 outcomes
                    # outcomes_r3 =~`3_ppvt_raw` +`3_maths_perco` +`3_yrs_grtr_grade` +
                                  # `3_literate` 
                # Round`4 outcomes
                    outcomes_ =~`4_ppvt_raw` +`4_maths_perco` +`4_yrs_grtr_grade`

              '
temp.model.fit <- cfa(temp.model, data = temp)
summary(temp.model.fit, fit.measures = T)
x <- modindices(temp.model.fit)
```


# 2  Now we look at structural + measurement model for R4 outcome

```{r struc+meas_r4}

temp <- peru_long.dat %>% 
          select(childid, round, typesite, indigenous,
                 mother_age_birth, partner, momedu, dadlit, momlit, 
                 wi_early_yrs, food_insecurity_r2, totalexp,
                 chore_hours, school_type, hours_study, read_encourage,
                 sdq1, sdq3, sdq4, sdq5, 
                 sdq9, sdq8, sdq10, sdq11, 
                 sdq13, sdq16, sdq12, sdq14,
                 sdq17, sdq20, sdq15, sdq18, 
                 sdq19,
                 drink_to_drunk, drunk_hit, alc_1perwk,
                 stunting_to_r3, probs_vision, probs_resp,
                 care_school_qual2, care_school_qual3,
                 care_school_qual4, 
                 school_punish1, school_punish2,
                 lang_instruct_matches,
                 ppvt_raw, maths_perco, yrs_grtr_grade, literate, deprived_grouping) %>% 
          melt(id.vars = c("childid","round")) %>% 
          dcast(childid ~ round + variable)

temp %<>% 
  mutate(`1_mother_age_birth` = `1_mother_age_birth`/25,
         `2_totalexp` = `2_totalexp`/750,
         `4_ppvt_raw`= `4_ppvt_raw`/200,
         `4_maths_perco`= `4_maths_perco`/200,
         `3_chore_hours`= `3_chore_hours`/2.2,
         `3_read_encourage` = `3_read_encourage`/3.2,
         `3_stunting_to_r3` = `3_stunting_to_r3`/2)

temp.model <- '
               # Caregiver Status
                   older_with_partner =~ `1_partner` 
                  # `1_mother_age_birth`
                   lit_ed =~`1_momedu` +`1_momlit`
                  
               # # Economic well-being
                    eco_well_being_cfa =~`2_wi_early_yrs` +`2_food_insecurity_r2` +`3_school_type`
                    # +`2_totalexp`

               # Education investment
                   ed_invest =~`3_chore_hours` +`3_hours_study`+
                                `3_read_encourage`
         

               # Adverse EC environment
                    mmhealth =~`1_sdq1` +`1_sdq3` +`1_sdq4` +`1_sdq5` + 
                               `1_sdq8` +`1_sdq10` + 
                               `1_sdq13` +`1_sdq14` +
                               `1_sdq17 `  +`1_sdq15` +`1_sdq18` +
                               `1_sdq19`
                    
                # Child health
                    child_health =~`3_stunting_to_r3` +`3_probs_vision` +`3_probs_resp`

                # School environment
                   #  school_qual =~`3_care_school_qual2` +`3_care_school_qual3` +
                   # `3_care_school_qual4` 
                    # school_punish =~`3_school_punish1` +`3_school_punish2`  
                     # lang_instruct =~`3_lang_instruct_matches`
                     # school_environ =~ school_qual + lang_instruct
                     # school_environ ~~ 1*school_environ

                # Round`3 outcomes
                    # outcomes_r3 =~`3_ppvt_raw` +`3_maths_perco` +`3_yrs_grtr_grade` +
                                  # `3_literate` 
                # Round`4 outcomes
                    outcomes_r4 =~`4_ppvt_raw` +`4_maths_perco` +`4_yrs_grtr_grade` 


                # REG COMPONENT
      
                     eco_well_being_cfa ~ lit_ed          
                     mmhealth ~ lit_ed 
                     child_health ~ eco_well_being_cfa 
                     ed_invest ~ eco_well_being_cfa  + lit_ed
                     # lang_instruct ~ eco_well_being_cfa
                     outcomes_r4 ~ ed_invest 
                      # + child_health + mmhealth + lang_instruct

              '

temp.model.fit <- sem(temp.model, data = temp)
summary(temp.model.fit, fit.measures = T)
inspect(temp.model.fit, "rsquare")
x <- modindices(temp.model.fit)
resid(temp.model.fit, type = "standardized")
summary(temp.model.fit, fit.measures = T, standardized = TRUE)

```


# 3  Now we look at structural + measurement model for R3

```{r struc+meas_r3}

# Let's repeat this for Round 3 outcomes
# But note that one has to load the original data again

temp <- peru_long.dat %>% 
          select(childid, round, typesite, indigenous,
                 mother_age_birth, partner, momedu, dadlit, momlit, 
                 wi_early_yrs, food_insecurity_r2, totalexp,
                 chore_hours, school_type, hours_study, read_encourage,
                 sdq1, sdq3, sdq4, sdq5, 
                 sdq9, sdq8, sdq10, sdq11, 
                 sdq13, sdq16, sdq12, sdq14,
                 sdq17, sdq20, sdq15, sdq18, 
                 sdq19,
                 drink_to_drunk, drunk_hit, alc_1perwk,
                 stunting_to_r3, probs_vision, probs_resp,
                 care_school_qual2, care_school_qual3,
                 care_school_qual4, 
                 school_punish1, school_punish2,
                 lang_instruct_matches,
                 ppvt_raw, maths_perco, yrs_grtr_grade, literate, deprived_grouping) %>% 
          melt(id.vars = c("childid","round")) %>% 
          dcast(childid ~ round + variable)

temp %<>% 
  mutate(`1_mother_age_birth` = `1_mother_age_birth`/25,
         `2_totalexp` = `2_totalexp`/750,
         `3_ppvt_raw`= `3_ppvt_raw`/200,
         `3_maths_perco`= `3_maths_perco`/200,
         `3_chore_hours`= `3_chore_hours`/2.2,
         `3_read_encourage` = `3_read_encourage`/3.2,
         `3_stunting_to_r3` = `3_stunting_to_r3`/2)

temp.model <- '
               # Caregiver Status
                   older_with_partner =~ `1_partner` 
                  # `1_mother_age_birth`
                   lit_ed =~`1_momedu` +`1_momlit`
                  
               # # Economic well-being
                    eco_well_being_cfa =~`2_wi_early_yrs` +`2_food_insecurity_r2` +`3_school_type`
                    # +`2_totalexp`

               # Education investment
                   ed_invest =~`3_chore_hours` +`3_hours_study`+
                                `3_read_encourage`
         

               # Adverse EC environment
                    mmhealth =~`1_sdq1` +`1_sdq3` +`1_sdq4` +`1_sdq5` + 
                               `1_sdq8` +`1_sdq10` + 
                               `1_sdq13` +`1_sdq14` +
                               `1_sdq17 `  +`1_sdq15` +`1_sdq18` +
                               `1_sdq19`
                    
                # Child health
                    child_health =~`3_stunting_to_r3` +`3_probs_vision` +`3_probs_resp`

                # School environment
                   #  school_qual =~`3_care_school_qual2` +`3_care_school_qual3` +
                   # `3_care_school_qual4` 
                    # school_punish =~`3_school_punish1` +`3_school_punish2`  
                     # lang_instruct =~`3_lang_instruct_matches`
                     # school_environ =~ school_qual + lang_instruct
                     # school_environ ~~ 1*school_environ

                # Round`3 outcomes
                   outcomes_r3 =~`3_ppvt_raw` +`3_maths_perco` +`3_yrs_grtr_grade` 
                                  # `3_literate` 
                # Round`4 outcomes
                    # outcomes_r4 =~`4_ppvt_raw` +`4_maths_perco` 


                # REG COMPONENT
      
                     eco_well_being_cfa ~ lit_ed          
                     mmhealth ~ lit_ed 
                     child_health ~ eco_well_being_cfa 
                     ed_invest ~ eco_well_being_cfa  + lit_ed
                     # lang_instruct ~ eco_well_being_cfa
                     outcomes_r3 ~ ed_invest 
                      # + child_health + mmhealth + lang_instruct

              '

temp.model.fit <- sem(temp.model, data = temp)
summary(temp.model.fit, fit.measures = T)
inspect(temp.model.fit, "rsquare")
x <- modindices(temp.model.fit)
resid(temp.model.fit, type = "standardized")
summary(temp.model.fit, fit.measures = T, standardized = TRUE)


```

# 4  Model by group : rural and indigenous vs urban and SPanish

```{r struc+meas_r4_rural2}

# Try to repeat model for disad group split

temp <- peru_long.dat %>% 
          select(childid, round,
                 mother_age_birth, partner, momedu, dadlit, momlit, 
                 wi_early_yrs, food_insecurity_r2, totalexp,
                 chore_hours, school_type, hours_study, read_encourage,
                 sdq1, sdq3, sdq4, sdq5, 
                 sdq9, sdq8, sdq10, sdq11, 
                 sdq13, sdq16, sdq12, sdq14,
                 sdq17, sdq20, sdq15, sdq18, 
                 sdq19,
                 drink_to_drunk, drunk_hit, alc_1perwk,
                 stunting_to_r3, probs_vision, probs_resp,
                 care_school_qual2, care_school_qual3,
                 care_school_qual4, 
                 school_punish1, school_punish2,
                 lang_instruct_matches,
                 ppvt_raw, maths_perco, yrs_grtr_grade, literate, deprived_grouping) %>% 
          melt(id.vars = c("childid","round")) %>% 
          dcast(childid ~ round + variable)
            
temp %<>% 
  mutate(`1_mother_age_birth` = `1_mother_age_birth`/25,
         `2_totalexp` = `2_totalexp`/750,
         `4_ppvt_raw`= `3_ppvt_raw`/200,
         `4_maths_perco`= `3_maths_perco`/200,
         `3_chore_hours`= `3_chore_hours`/2.2,
         `3_read_encourage` = `3_read_encourage`/3.2,
         `3_stunting_to_r3` = `3_stunting_to_r3`/2)

temp.model <- '
               # Caregiver Status
                   # older_with_partner =~ `1_partner`
                        # + `1_mother_age_birth`
                   lit_ed =~`1_momedu` +`1_momlit`
                  
               # # Economic well-being
                    eco_well_being_cfa =~`2_wi_early_yrs` +`2_food_insecurity_r2` 
                    # +`2_totalexp`

               # Education investment
                   ed_invest =~`3_chore_hours` +`3_hours_study`+
                                `3_read_encourage`
         

               # Adverse EC environment
                    mmhealth =~`1_sdq1` +`1_sdq3` +`1_sdq4` +`1_sdq5` + 
                               `1_sdq8` +`1_sdq10` + 
                               `1_sdq13` +`1_sdq14` +
                               `1_sdq17 `  +`1_sdq15` 
                               
                    
                # Child health
                    child_health =~`3_stunting_to_r3` 
                                    # +`3_probs_vision` 
                                    # +`3_probs_resp`

                # School environment
                   #  school_qual =~`3_care_school_qual2` +`3_care_school_qual3` +
                   # `3_care_school_qual4` 
                    # school_punish =~`3_school_punish1` +`3_school_punish2`  
                     # lang_instruct =~`3_lang_instruct_matches`
                     # school_environ =~ school_qual + lang_instruct
                     # school_environ ~~ 1*school_environ

                # Round`3 outcomes
                    # outcomes_r3 =~`3_ppvt_raw` +`3_maths_perco` +`3_yrs_grtr_grade` +
                                   # `3_literate` 
                # Round`4 outcomes
                    outcomes_r4 =~`4_ppvt_raw` +`4_maths_perco` 


                # REG COMPONENT
      
                     eco_well_being_cfa ~ lit_ed          
                     mmhealth ~ lit_ed 
                     child_health ~ eco_well_being_cfa 
                     ed_invest ~  mmhealth + eco_well_being_cfa
                     # lang_instruct ~ eco_well_being_cfa
                     outcomes_r4 ~ ed_invest + eco_well_being_cfa 
                      # + child_health + lang_instruct

              '

temp$depriv_group = temp$`1_deprived_grouping`
temp %<>% 
  filter(depriv_group == 1)

temp.model.fit <- sem(temp.model, data = temp)
summary(temp.model.fit, fit.measures = T)
inspect(temp.model.fit, "rsquare")
x <- modindices(temp.model.fit)
summary(temp.model.fit, fit.measures = T, standardized = TRUE)


```


# 5  Model by group : boys vs girls

```{r struc+meas_r4_sex}

# Try to repeat model for disad group split

temp <- peru_long.dat %>% 
          select(childid, round, typesite, indigenous, sex,
                 mother_age_birth, partner, momedu, dadlit, momlit, 
                 wi_early_yrs, food_insecurity_r2, totalexp,
                 chore_hours, school_type, hours_study, read_encourage,
                 sdq1, sdq3, sdq4, sdq5, 
                 sdq9, sdq8, sdq10, sdq11, 
                 sdq13, sdq16, sdq12, sdq14,
                 sdq17, sdq20, sdq15, sdq18, 
                 sdq19,
                 drink_to_drunk, drunk_hit, alc_1perwk,
                 stunting_to_r3, probs_vision, probs_resp,
                 care_school_qual2, care_school_qual3,
                 care_school_qual4, 
                 school_punish1, school_punish2,
                 lang_instruct_matches,
                 ppvt_raw, maths_perco, yrs_grtr_grade, literate, deprived_grouping) %>% 
          melt(id.vars = c("childid","round")) %>% 
          dcast(childid ~ round + variable)

temp %<>% 
  mutate(`1_mother_age_birth` = `1_mother_age_birth`/25,
         `2_totalexp` = `2_totalexp`/750,
         `4_ppvt_raw`= `4_ppvt_raw`/200,
         `4_maths_perco`= `4_maths_perco`/200,
         `3_chore_hours`= `3_chore_hours`/2.2,
         `3_read_encourage` = `3_read_encourage`/3.2,
         `3_stunting_to_r3` = `3_stunting_to_r3`/2)

temp.model <- '
               # Caregiver Status
                   older_with_partner =~ `1_partner` 
                  # `1_mother_age_birth`
                   lit_ed =~`1_momedu` +`1_momlit`
                  
               # # Economic well-being
                    eco_well_being_cfa =~`2_wi_early_yrs` +`2_food_insecurity_r2` +`3_school_type`
                    # +`2_totalexp`

               # Education investment
                   ed_invest =~`3_chore_hours` +`3_hours_study`+
                                `3_read_encourage`
         

               # Adverse EC environment
                    mmhealth =~`1_sdq1` +`1_sdq3` +`1_sdq4` +`1_sdq5` + 
                               `1_sdq8` +`1_sdq10` + 
                               `1_sdq13` +`1_sdq14` +
                               `1_sdq17 `  +`1_sdq15` +`1_sdq18` +
                               `1_sdq19`
                    
                # Child health
                    child_health =~`3_stunting_to_r3` +`3_probs_vision` +`3_probs_resp`

                # School environment
                   #  school_qual =~`3_care_school_qual2` +`3_care_school_qual3` +
                   # `3_care_school_qual4` 
                    # school_punish =~`3_school_punish1` +`3_school_punish2`  
                     # lang_instruct =~`3_lang_instruct_matches`
                     # school_environ =~ school_qual + lang_instruct
                     # school_environ ~~ 1*school_environ

                # Round`3 outcomes
                    # outcomes_r3 =~`3_ppvt_raw` +`3_maths_perco` +`3_yrs_grtr_grade` +
                                  # `3_literate` 
                # Round`4 outcomes
                    outcomes_r4 =~`4_ppvt_raw` +`4_maths_perco` + `4_yrs_grtr_grade`


                # REG COMPONENT
      
                     eco_well_being_cfa ~ lit_ed          
                     mmhealth ~ lit_ed 
                     child_health ~ eco_well_being_cfa 
                     ed_invest ~ eco_well_being_cfa  + lit_ed
                     # lang_instruct ~ eco_well_being_cfa
                     outcomes_r4 ~ ed_invest 
                      # + child_health + mmhealth + lang_instruct

              '
temp$sex <- temp$`1_sex`

temp.model.fit <- sem(temp.model, data = temp, group = "sex")
summary(temp.model.fit, fit.measures = T)
inspect(temp.model.fit, "rsquare")
x <- modindices(temp.model.fit)
resid(temp.model.fit, type = "standardized")
summary(temp.model.fit, fit.measures = T, standardized = TRUE)


```

